---
title: "Data frame manipulations base R"
---

## Data Frame

In this tutorial, we show how to create a data frame. Then we move on to some manipulations with data frames.

It is one of the most frequent use objects in R. It is not atomic - it can contain more than one data types. For instance, some variables may be characters and others numeric.

A data frame object of not recursive. That is, a data frame cannot contain another data frame.

A data frame object has attributes: names, row.names, and class.

## Creating a Data Frame

```{r}
data_frame <- data.frame(id = c(1,2,3,4), income = rnorm(4, 100, 2), 
                         expense = rnorm(4, 50, 1))

summary(data_frame)

second_data_frame <- data.frame( id = c(2,3,6,7), wages = rnorm(4, 20, 1))


```

## Sorting Data Frame

Sorting a data frame is an operation that is often done. One way to do it is to use the function "order()" which will return a vector of indices of the sorted data.frame.

```{r}
data_frame_sorted <- data_frame[order(data_frame$income, decreasing = TRUE),]
data_frame_sorted
```

## Summary Statistics from Data Frame

```{r}
aggregate(x = state.x77, by = list(Region = state.region), mean)

```

```{r}
testDF <- data.frame(v1 = c(1,3,5,7,8,3,5,NA,4,5,7,9),
                     v2 = c(11,33,55,77,88,33,55,NA,44,55,77,99) )

by1 <- c("red", "blue", 1, 2, NA, "big", 1, 2, "red", 1, NA, 12)
by2 <- by2 <- c("wet", "dry", 99, 95, NA, "damp", 95, 99, "red", 99, NA, NA)

aggregate(x = testDF, by = list(by1), FUN = "mean")
```

```{r}
# produce the average of weight by feed.
aggregate(weight ~ feed, data = chickwts, mean)
  
```

Using a formula with more than two other variables.

```{r}
warpbreaks
aggregate(breaks ~ wool + tension, data = warpbreaks, mean)
```

```{r}
airquality
aggregate(cbind(Ozone, Temp) ~ Month, data = airquality, mean)
```
