---
title: "Factors"
---

## Factor

Factor is an object in R usually uses for the manipulations of categorical variables. Factor is used to encode a vector as a factor (the terms 'category' and 'enumerated type' are used for factors).

## Creating a factor variable

```{r}
# Create a vector with some categorical data
data_vector <- c("A", "B", "A", "C", "B", "A")

# Create a factor variable
factor_variable <- factor(data_vector)

# Print the factor variable
print(factor_variable)


```

## Specifying Custom Levels

You can custom levels for your factor variable using the 'levels' argument:

```{r}

# Create a factor variable with custom levels
custom_levels <- factor(data_vector, levels = c("A","B","C"))


# Print the factor variable with custome levels
print(custom_levels)
```

## Changing the Order of levels

The order of levels in a factor can be changed using the 'levels' function.

```{r}
factor_variable <- factor(data_vector, levels = c("C","A","B"))

# Print the factor variable with changed levels
print(factor_variable)
```

## Summary Statistics

Factor variables are useful for summarizing data. Functions such as 'table' can be used in combination with factor variables.

```{r}
summary_variable <- table(factor_variable)
print(summary_variable)
```

## Using Factor in Statistical Analysis

```{r}
# Create a numeric variable
numeric_data <- c(10, 15, 12, 20, 18, 25)

factor_variable <- factor(data_vector, levels = c("A","B"))

# Perform a simple t-test by factor level
t_test_result <- t.test(numeric_data ~ factor_variable)

# Print the t-test result
print(t_test_result)

```
