---
title: "Lists"
---

## List
A list is a versatile data structure in R that can hold elements of different data types.

Lists are not atomic; they are composite data structures. 

Lists can be recursive, meaning they can contain other lists.

Lists can have attributes, which are metadata associated with the list. These attributes can include names, dimensions, or other information.

## Create

```{r}
# Create an empty list
my_list <- list()

# Create a list with elements of different types
mixed_list <- list(1, "apple", 3.14, TRUE)

# Create a list with named elements
named_list <- list(name = "Alice", age = 30, city = "New York")

```


### Attributes
```{r}
attributes(mixed_list)

# Adding names to a list
names(mixed_list) <- c("Number", "Fruit", "Value", "Boolean")

# Adding dimensions to a list
dim(mixed_list) <- c(2, 2)

attributes(mixed_list)

```


## Read

```{r}
# Access elements by index
first_element <- mixed_list[[1]]
second_element <- mixed_list[[2]]

# Access elements by name
alice_age <- named_list$name
alice_city <- named_list$city

```

## Update

```{r}
# Modify an element
mixed_list[[2]] <- "banana"

# Add a new element
mixed_list[[5]] <- "cherry"

# Modify named elements
named_list$age <- 31
named_list$city <- "Los Angeles"

```

## Delete

```{r}
# Remove an element by index
rm_element <- mixed_list[[3]]
mixed_list <- mixed_list[-3]

# Remove an element by name
rm(named_list)

```

